\ourlongtitle

### 2021-08-27 Bernoulli/AnswerReviews.tex

We would like to thank the associate editor and the referees for the attention they paid to our work and for their positive feedbacks.

\bigskip

Minor errors and typos listed by R2 were corrected as suggested. Please find below a few detailed answers.

\begin{itemize}
\item It might have made the paper less scary and dry to read to first provide standard setups (along the lines of what is at the top of p.2), then proceed to the objective (formula (1.2)) and only then supplement all the measure-theoretic setup, e.g. as part of section 2. This would also have yielded the advantage of collecting mathematical notation at the start of section 2 (Is a collection of common notation in tabular form allowed in Bernoulli? This would be a friendly service to the reader.) rather than scattering it over the start of section 1 and section 2.\\
{\em We thank the referee for this suggestion. We postponed the introduction of (1.2)  and the objective of the paper (additive smoothing) after the standard setups (Feynman-Kac path models and hidden Markov models) for better clarity.}
\item p.7 (2.10) and elsewhere uses rectangular brackets as in $\phi_m[\ell_m(\cdot,x_{m+1})]$ which I think is meant to denote $\int_{X_m}\ell_m(x_m,x_{m+1})\phi_m(dx_m)$. This notation has not been introduced. The "Preliminaries" at the top of p.4 have $\mu h$ for the integral of a function $h$ against a measure $\mu$ which would suggest $\phi_m \ell(\cdot,x_{m+1})$. Since the rectangular brackets are used in this sense at many points in the paper, perhaps this notation should be added to the preliminaries.\\
{\em The notation is now introduced at the beginning of Section~2. }
\item p.7 Just before 2.2.2, it would be nice to comment on where one might get suitable adjustment weights $\vartheta_n$ from. How are these to be chosen or tuned?\\
{\em We added the following lines after the introduction of the adjustment-weight function (after (2.8): "Adjustment multiplier weights were introduced in [31] to propose data-driven proposal distributions and design robust filtering estimates. These weights usually depend on the new observation and are often chosen to approximate the predictive likelihood of the new observation given the current particle. For further discussions see for instance [7,Chapter~8]".}
\item p. 15-16: for the presentation of Example 2, I feel it would be better to describe the problem fully before giving Durham and Gallant's solution.\\
{\em In Example~2, we thought it would be better to first fully describe the transition density estimator of the process defined by (3.8) before discussing several applications depending on the chosen model for the observation process.}
\item p.29 Appendix A in the supplement has $\phi_n(dx_n)[\ell(\cdot,x_{n+1})]$  in the numerator that arises from $B_n(x_{n+1},dx_n)$. I think this should read $\phi_n(dx_n)\ell(x_n,x_{n+1})$  instead?\\
{\em The normalizing term given in the appendix is correct, see (2.10).}
\item p.30 Appendix A has $\phi_{0:n}L_{0,n}1_{X^n}$  in the denominator for  $\phi_{0:n+1}h$ which I think should read  $\phi_{0:n}L_{0,n}1_{X^{n+1}}$ as there is a dangling $dx_{n+1}$ otherwise.\\
{\em There was indeed a mistake in this identity, thank you for pointing this out. This was corrected using (2.3).}
\end{itemize}

### 2021-08-27 Bernoulli/glo-2020-theor-results.tex
Final corrections

To be able to describe our results in full detail, we also need to introduce, for every $n \in \nset$, the skew backward Markov kernels 
$$
    \bkmod{n}(x_{n + 1}, \rmd x_n) \eqdef \frac{\postmod{n}(\rmd x_n) \, \udmod{n}(x_n, x_{n + 1})}{\postmod{n}[\udmod{n}(\cdot, x_{n + 1})]}
$$
on $\Xset_{n + 1} \times \Xfd_n$ as well as the joint law 
\begin{equation} \label{eq:skew:backward:law}
\tstatmod{n}(x_n, \rmd x_{0:n - 1}) \eqdef \prod_{m = 0}^{n - 1} \bkmod{m}(x_{m + 1}, \rmd x_m)
\end{equation}
on $\Xset_n \times \Xfd^{n - 1}$.

The analysis will be carried through under the following assumption.
\begin{hypH}
\label{assum:bound:filter:pseudomarginal}
For all $n \in \nset$, the functions $\adjfuncforward{n}$ and  
\begin{align*}
&\wgtfuncext{n} : \Xset_n \times \Xset_{n + 1} \times \Zset_{n + 1} \ni (x_n, x_{n + 1}, z) \mapsto \frac{\ukest{n}{z}(x_n, x_{n + 1})}{\adjfuncforward{n}(x) \kissforward{n}{n}(x_n, x_{n + 1})}, \\
&\wgtfuncmod{n} : \Xset_n \times \Xset_{n + 1} \ni (x_n, x_{n + 1}) \mapsto \frac{\udmod{n}(x_n, x_{n + 1})}{\adjfuncforward{n}(x_n) \kissforward{n}{n}(x_n, x_{n + 1})}
\end{align*}
are bounded. So is also $\initwgtfunc : \Xset_0 \ni x_0 \mapsto \rmd \chi / \rmd \init(x_0)$. 
\end{hypH}

@@

Our first theoretical result is the following Hoeffding-type concentration inequality, which also plays a critical role for the derivation of the central limit theorem (CLT) in Theorem~\ref{cor:clt:pseudo:marginal:paris}. For every $n \in \nset$, let $\bmaf{\Xfd^n}$ denote the set of additive functionals \eqref{eq:add:func} such that $\addf{m} \in \bmf{\Xfd_m \tensprod \Xfd_{m + 1}}$ for all $m \in \intvect{0}{n - 1}$.   
\begin{theorem}
\label{cor:hoeffding:tau:marginal}
Assume \hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for every $n \in \nset$, $\precpar \in \precparsp$, $h_n \in \bmaf{\Xfd^n}$, and $\K \in \nsetpos$ there exists $(c_n, d_n) \in \rsetpos^2$ such that for all $\N \in \nsetpos$ and $\delta > 0$,
$$
\pP\left(\left| \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n \right| \geq \delta \right) \leq c_n \exp \left( - d_n \N \delta^2 \right).
$$
\end{theorem}

@@

We now focus on the asymptotic properties of Algorithm~\ref{alg:pm:PaRIS} and furnish, in Theorem~\ref{cor:clt:pseudo:marginal:paris} below, this scheme with a CLT. On the basis of this result, the stochastic stability of the algorithm is expressed by establishing, in Theorem~\ref{thm:variance:bound}, an $\mathcal{O}(n)$ bound on the asymptotic variances. In order to be able to state these results accurately, we need some additional notation. First, let 
 for $(x_n, x_{n + 1}) \in \Xset_n \times \Xset_{n + 1}$, 
\begin{multline} \label{eq:def:ukestvar}
\ukestvar{n}(x_n, x_{n + 1}) \eqdef 
\frac{1}{\wgtfuncmod{n}(x_n, x_{n + 1})} \int \{ \wgtfunc{n}(x_n, x_{n + 1}, z) - \wgtfuncmod{n}(x_n, x_{n + 1}) \}^2 \, \ukdist{n}(x_n, x_{n + 1}, \rmd z)
\end{multline}
denote the conditional relative variance of the (random) weight $\ewght{n + 1}{i}$ given $\epart{n}{\ind{n}{i}} = x_n$ and $\epart{n + 1}{i}= x_{n + 1}$. In addition, for each $n \in \nset$ and $m \in \intvect{0}{n}$, define the kernel  
\begin{equation} \label{eq:def:retrokmodmod}
    \retrokmodmod_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstatmod{m}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \ukmod{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
\begin{equation} \label{eq:def:retrokmodmodnorm}
\retrokmodmodnorm_{m, n} h (x_m) \eqdef  \retrokmodmod_{m, n}(h - \postmod{0:n} h)(x_m) 
\end{equation}
on the same space.

The following is the main result of this section. 
\begin{theorem} \label{cor:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n  \right) 
  \dlim \sigma_n(h_n) Z, 
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:as:var:decomp}
\sigma^2_n (h_n) \eqdef \frac{\chi(\initwgtfunc \retrokmodmodnorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) + \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) 
\end{equation}
and
\begin{multline} \label{eq:non-recursive:as:var}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\retrokmodmodnorm_{m + 1, n} h_n ]^2)}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}\\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{theorem}

>>>>

To be able to describe our results in full detail, we also need to introduce, for every $n \in \nset$, the skew backward Markov kernels 
$$
    \bkmod{n}(x_{n + 1}, \rmd x_n) \eqdef \frac{\postmod{n}(\rmd x_n) \, \udmod{n}(x_n, x_{n + 1})}{\postmod{n}[\udmod{n}(\cdot, x_{n + 1})]}
$$
on $\Xset_{n + 1} \times \Xfd_n$ as well as the joint law 
\begin{equation} \label{eq:skew:backward:law}
\tstatmod{n}(x_n, \rmd x_{0:n - 1}) \eqdef \prod_{m = 0}^{n - 1} \bkmod{m}(x_{m + 1}, \rmd x_m)
\end{equation}
on $\Xset_n \times \Xfd^{n - 1}$. The analysis will be carried through under the following assumption.
\begin{hypH}
\label{assum:bound:filter:pseudomarginal}
For all $n \in \nset$, the functions $\adjfuncforward{n}$ and  
\begin{align*}
&\wgtfuncext{n} : \Xset_n \times \Xset_{n + 1} \times \Zset_{n + 1} \ni (x_n, x_{n + 1}, z) \mapsto \frac{\ukest{n}{z}(x_n, x_{n + 1})}{\adjfuncforward{n}(x) \kissforward{n}{n}(x_n, x_{n + 1})}, \\
&\wgtfuncmod{n} : \Xset_n \times \Xset_{n + 1} \ni (x_n, x_{n + 1}) \mapsto \frac{\udmod{n}(x_n, x_{n + 1})}{\adjfuncforward{n}(x_n) \kissforward{n}{n}(x_n, x_{n + 1})}
\end{align*}
are bounded. So is also $\initwgtfunc : \Xset_0 \ni x_0 \mapsto \rmd \chi / \rmd \init(x_0)$. 
\end{hypH}

@@

Our first result is the following Hoeffding-type concentration inequality, which also plays a critical role for the derivation of the central limit theorem (CLT) in Theorem~\ref{cor:clt:pseudo:marginal:paris}. For every $n \in \nset$, let $\bmaf{\Xfd^n}$ denote the set of additive functionals \eqref{eq:add:func} such that $\addf{m} \in \bmf{\Xfd_m \tensprod \Xfd_{m + 1}}$ for all $m \in \intvect{0}{n - 1}$.   
\begin{theorem}
\label{cor:hoeffding:tau:marginal}
Assume \hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for every $n \in \nset$, $\precpar \in \precparsp$, $h_n \in \bmaf{\Xfd^n}$, and $\K \in \nsetpos$ there exists $(c_n, d_n) \in \rsetpos^2$ such that for all $\N \in \nsetpos$ and $\delta > 0$,
$$
\pP\left(\left| \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n \right| \geq \delta \right) \leq c_n \exp \left( - d_n \N \delta^2 \right).
$$
\end{theorem}

@@

We now focus on the asymptotic properties of Algorithm~\ref{alg:pm:PaRIS} and furnish, in Theorem~\ref{cor:clt:pseudo:marginal:paris} below, this scheme with a CLT. On the basis of this result, the stochastic stability of the algorithm is expressed by establishing, in Theorem~\ref{thm:variance:bound}, an $\mathcal{O}(n)$ bound on the asymptotic variances. In order to be able to state these results accurately, we need some additional notation. First, let 
 for $(x_n, x_{n + 1}) \in \Xset_n \times \Xset_{n + 1}$, 
\begin{multline} \label{eq:def:ukestvar}
\ukestvar{n}(x_n, x_{n + 1}) \eqdef 
\frac{1}{\wgtfuncmod{n}(x_n, x_{n + 1})} \int \{ \wgtfunc{n}(x_n, x_{n + 1}, z) - \wgtfuncmod{n}(x_n, x_{n + 1}) \}^2 \, \ukdist{n}(x_n, x_{n + 1}, \rmd z)
\end{multline}
denote the conditional relative variance of the (random) weight $\ewght{n + 1}{i}$ given $\epart{n}{\ind{n}{i}} = x_n$ and $\epart{n + 1}{i}= x_{n + 1}$. In addition, for each $n \in \nset$ and $m \in \intvect{0}{n}$, define the kernel  
\begin{equation} \label{eq:def:retrokmodmod}
    \retrokmodmod_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstatmod{m}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \ukmod{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
\begin{equation} \label{eq:def:retrokmodmodnorm}
\retrokmodmodnorm_{m, n} h (x_m) \eqdef  \retrokmodmod_{m, n}(h - \postmod{0:n} h)(x_m) 
\end{equation}
on the same space. The following is the main result of this section. 
\begin{theorem} \label{cor:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n  \right) 
  \dlim \sigma_n(h_n) Z, 
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:as:var:decomp}
\sigma^2_n (h_n) \eqdef \frac{\chi(\initwgtfunc \retrokmodmodnorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) + \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) 
\end{equation}
and
\begin{multline} \label{eq:non-recursive:as:var}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\retrokmodmodnorm_{m + 1, n} h_n ]^2)}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}\\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{theorem}
